### 2019-01-28 2012.fall.pku/project02.tex
Add files via upload

\lecture{Project I}{Yuan Yao}{Due: Monday Oct 8, 2012}{Sep 15, 2012}

The first project aims to have more experiments on PCA/MDS. Team work is encouraged, but you have to \emph{write your proposal yourself} with a full citation of partners' work. Submit to TA your report with your source codes (as appendix in report or zipped files).

On course website, you will find the following datasets

@@

@@

\url{http://www.math.pku.edu.cn/teachers/yaoy/Fall2011/xiyouji/xiyouji.mat}

@@

\url{http://www.math.pku.edu.cn/teachers/yaoy/Fall2011/xiyouji/readData.m}

@@

\url{http://www-stat.stanford.edu/\~{}tibs/ElemStatLearn/datasets/zip.digits/}

@@

@@

Pick up your favorite dataset(s), perform PCA or MDS to analyze and visualize the data, with denoising and outliers removal if necessary. Attention: the last dataset is relatively big with about 2GB size.

>>>>

\lecture{Mini-Project II}{Yuan Yao}{Due: Monday Nov 19, 2012}{Nov 5, 2012}

This project aims to exercise the tools in the class, such as random projections, robust PCA, sparse PCA, and MDS with uncertainty, etc., based on the real datasets. In the below, we list 6 candidate datasets for your reference. 
\begin{enumerate}
%\item Choose one or more datasets, either from those listed in the first project or by your own. Write in your report what kind of \emph{problems} you would like to attack with the data. (It is even more important to find a proper problem than solving it eventually!)
\item Choose the dataset in your favorite. In your report, raise a problem that you are interested to attack with the tools you learned. The following problems are for your examples:
\subitem  Robust PCA will decompose the data matrices into two parts, a low-rank and a sparse, which can be applied to all the datasets below. What's the meaning of sparse part and what's the meaning of low-rank part in your data? 
\subitem Sparse PCA will find PCA whose support are only subsets of the variables. Journey to the West and Chinese Medicine datasets are probably suitable for such a tool, as we would like to find small teams of characters dominating the story, or small sets of essential herb curing most of similar deseases (flu). 
\subitem Random Projection is to do dimensionality reduction if one wants to find SVD of $X$ of large dimensionality. This particularly suits the analysis of SNPs data with $650k$ sites, where you can random projects the matrix into a subset of SNPs and see if top eigenvectors are still approximately correct. 
\subitem MDS with uncertainty is to reconstruct Euclidean coordinates when pairwise distances are partially observed and contaminated.  Can one reconstruct Euclidean coordinates of Bird Flu virus from geodesic distances and sequence distances, respectively, and compare one against another?  
\item Try to apply the methods to analyze your data. Analyze the phenomena you observed and explain whether or not the results answer your problems in the first step and \emph{why}. Finally in your report list the kind of \emph{open problems} in your project and possible future directions in your mind. 
\end{enumerate}

\section*{Datasets}

@@

\item {\em Bird Flu Dataset:} (courtesy of Steve Smale and Cissy) This dataset 162 H5N1 (bird flu) virus sequences discovered around the world:

\url{http://www.math.pku.edu.cn/teachers/yaoy/data/birdflu_seq162.txt}

Locations of such virus discovered are reported with latitude and longitude coordinates on the globe:

\url{http://www.math.pku.edu.cn/teachers/yaoy/data/birdflu_latgrat.txt}

Pairwise geodesic distances between these 162 sites are constructed as

\url{http://www.math.pku.edu.cn/teachers/yaoy/data/birdflu_geodist.txt}

A kernel-induced $l_2$-distances between 162 virus sequences are given in

\url{http://www.math.pku.edu.cn/teachers/yaoy/data/birdflu_l2dist.txt}

\item {\em Chinese Medicine Dataset:} (courtesy of Zhi Geng) This dataset contains a 2885-by-965 matrix $X$ with 2885 diseases (mostly flu) and 965 Chinese herbal prescription drugs collected from traditional Chinese medicine literature. Each element contains either $0$ or $1$ indicating if the herbal drug is used for the disease. Variable {\tt{desease2885}} is a 1-by-2885 cell collecting the description of deseases and {\tt{herb965}} contains the names of those herbal drugs.

\url{http://www.math.pku.edu.cn/teachers/yaoy/data/cmed965.mat}

@@

\url{http://www.math.pku.edu.cn/teachers/yaoy/data/xiyouji/xiyouji.mat}

@@

\url{http://www.math.pku.edu.cn/teachers/yaoy/data/xiyouji/readData.m}

@@

\url{http://www-stat.stanford.edu/\~tibs/ElemStatLearn/datasets/zip.digits/}

@@

Attention: this last dataset is relatively big with about 2GB size.

@@

### 2019-01-28 2012.fall.pku/project_final.tex
Add files via upload

\lecture{Mini-Project II}{Yuan Yao}{Due: Monday Nov 19, 2012}{Nov 5, 2012}

This project aims to exercise the tools in the class, such as random projections, robust PCA, sparse PCA, and MDS with uncertainty, etc., based on the real datasets. In the below, we list 6 candidate datasets for your reference. 
\begin{enumerate}
%\item Choose one or more datasets, either from those listed in the first project or by your own. Write in your report what kind of \emph{problems} you would like to attack with the data. (It is even more important to find a proper problem than solving it eventually!)
\item Choose the dataset in your favorite. In your report, raise a problem that you are interested to attack with the tools you learned. The following problems are for your examples:
\subitem  Robust PCA will decompose the data matrices into two parts, a low-rank and a sparse, which can be applied to all the datasets below. What's the meaning of sparse part and what's the meaning of low-rank part in your data? 
\subitem Sparse PCA will find PCA whose support are only subsets of the variables. Journey to the West and Chinese Medicine datasets are probably suitable for such a tool, as we would like to find small teams of characters dominating the story, or small sets of essential herb curing most of similar deseases (flu). 
\subitem Random Projection is to do dimensionality reduction if one wants to find SVD of $X$ of large dimensionality. This particularly suits the analysis of SNPs data with $650k$ sites, where you can random projects the matrix into a subset of SNPs and see if top eigenvectors are still approximately correct. 
\subitem MDS with uncertainty is to reconstruct Euclidean coordinates when pairwise distances are partially observed and contaminated.  Can one reconstruct Euclidean coordinates of Bird Flu virus from geodesic distances and sequence distances, respectively, and compare one against another?  
\item Try to apply the methods to analyze your data. Analyze the phenomena you observed and explain whether or not the results answer your problems in the first step and \emph{why}. Finally in your report list the kind of \emph{open problems} in your project and possible future directions in your mind. 
\end{enumerate}

\section*{Datasets}

>>>>

\lecture{Final Project}{Yuan Yao}{Due: Thursday Jan 3rd, 2013}{December 20, 2012}

The final project encourages students to work in small team, but write the final report \emph{independently}. Teammates may share their data, methodology, and results. But in the report, you have to 
\begin{itemize}
\item \emph{make your own analysis};
\item \emph{give proper credit to the work of your partner}. 
\end{itemize}

%\item Choose one or more datasets, either from those listed in the first project or by your own. Write in your report what kind of \emph{problems} you would like to attack with the data. (It is even more important to find a proper problem than solving it eventually!)
\section{Topic Models} 
You can login my server:

\texttt{ssh einstein@162.105.68.237}

\noindent using the password I provided on class. Some data are provided there:

\begin{itemize}
\item \texttt{/data/twitter7/}   contains 
\subitem June-Dec 2009 tweets, in \texttt{tweets2009-**.txt} files
\subitem \texttt{twitter\_rv.net}  a who-follow-whom network ($9M$), whose id-name correspondence is in file \texttt{numeric2screen}
\item \texttt{/home/einstein/ZHAOXIN/TweetMerged}, Zhao Xin's Singapore tweet data, one line consists of all the tweets of a user
\item \texttt{/data/nytimes/}   contains 50k articles about China in NewYork Times since 1989
\item \texttt{/data/wikipedia/dbpedia}   contains wikipedia articles
\end{itemize}

The following softwares are helpful 
\begin{itemize}
\item \texttt{/home/einstein/ZHAOXIN/GibbsLDA++-0.2/src}, whose usage can be found at \url{http://gibbslda.sourceforge.net/}
\item if you like matlab, you may try 
\url{http://psiexp.ss.uci.edu/research/programs_data/toolbox.htm}
\end{itemize}

\section{Big Matrix Factorization}

For those who are interested to do matrix factorization (SVD etc.) with big matrices, the following paper

Divide-and-Conquer Matrix Factorization, by Lester Mackey, Ameet Talwalkar, Michael I. Jordan, version 6, Aug 2012. 
\url{http://arxiv.org/abs/1107.0789v6}

\noindent summarizes two important class of methods:
\begin{itemize}
\item random projection (including column projection)
\item subsampling as generalized Nystr\"{o}m Method for SVD
\end{itemize}

You may implement one or two of these methods on your problem and test their efficiency.

\section{Chess Player Rating}

For those interested in Chess player rating in the Kaggle competition, the following data and algorithm are for your reference 
\begin{itemize}
\item \texttt{/data/chess/data/} contains the following data
\subitem \texttt{training\_data.csv}: training data
\subitem \texttt{cross\_validation\_dataset.csv}  
\subitem \texttt{test\_scores.csv}: test data with game scores
\subitem \texttt{players.csv} contains 8k chess players in the data
\item \texttt{/data/chess/R/}  contains two algorithms
\subitem \texttt{elopp.R}  is the kaggle award winner algorithm (MSRE=0.69), whose reference can be found at 
\subitem \texttt{/data/chess/Reference/}, or the paper:  
How I won the "Chess Ratings - Elo vs the Rest of the World" Competition, by Yannis Sismanis, 2010, \url{http://arxiv.org/abs/1012.4571}
\subitem \texttt{hodgeRank1.R} is a prototype algorithm for hodgeRank by Dr. Quanwu Xiao.
\end{itemize}

\section{Other topics}

You can choose other datasets for the following tasks

\begin{itemize}
\item dimensionality reduction and estimation (PCA, MDS, ISOMAP, LLE, Diffusion, Laplacian, LTSA)
\item spectral clustering 
\item RPCA and SPCA  
\end{itemize}

In these tasks, explore the ``criteria" to justify whether your method and choice of parameters is good or not.

\section{Paper Reading Project}

If you prefer to a solid paper reading project, please email me with your interested paper (if not listed on the class) or general interests. A paper presentation will be assigned to you.

\section*{Old Datasets}
